Efficient Parallelization of $\ell_1$-Minimization for Face Recognition \thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}

### 2011-02-09 paper/main.tex
Moved figure down to where it goes.
Restructured paper

\section{Augmented Lagrangian Multiplier Algorithm}
%Describe the two variants of the ALM algorithm

@@

%Show Pseudocode for one variant
%\subsection{Computational Complexity}
%The Matrix-vector computions are O($mn$), and the vector-vector operations are O($m$),  in terms of both computation cost and memory-bandwidth.  As A'A can be precomputed once, assuming the iterative solver always takes a constant number of iterations, the PALM algorithm has a computational complexity of O($mn+n^2$), where n $<$ m.

% Complexity of entire algorithm

\subsection{Adapations}

In this section, we describe how to adapt the general ALM algorithm for the alignment and recognition problem.  Refer to book...

In order to solve the alignment stage before recognition, we consider the $l_1$ problem: 
\begin{equation} 
\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
\end{equation}
which is solved in a similar fashion.

\section{CPU Implementation}
% With a reshuffling of the order of operations and a change of the termination
% condition, the inner loop takes the form of two matrix-vector multiplications
% and a sequence of vector-vector operations.

% put this computational complexity in section 2?
\subsection{Computational Complexity} Any discussion of the speed of an
algorithm should discuss the comutational complexity of the algorithm.  For the
ALM algorithms considered in this paper, this is rather simple.  The
Matrix-vector computions are $O(mn)$, and the vector-vector operations are
$O(m)$, in terms of both computation cost and memory-bandwidth.  Since A is
very tall, the computation of $A^\dagger$, the pseudoinverse of $A$, is best
computed via $A^\dagger = (A^TA)^{-1} A^T$.  The complexity of computing $G =
A^T A$ is $O(nnm)$.  $G$ is sufficiently small that it can be inverted via
Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation
of $A$ finishes with a matrix multiplication, with cost $O(n^2 m)$.
Fortunately, $A^\dagger$ only needs to be computed once per instance of the
$\ell_1$ minimization problem.

@@

\section{Hardware Parallelism}

@@

%\subsection{Solving a Single $\ell_1$ minimization}

@@

% This section is Victor's 
\section{Single Instance Implementations}

\subsection{Single Instance CPU Implementations}

\subsection{Single Instance GPU Implementations}
The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
CUBLAS are the BLAS libraries consisting of all single precision BLAS calls
which are provided by Nvidia and are optimized for GPU performance. In
addition, we coded kernel (GPU) methods for vector-vector operations to replace
operations that would have taken multiple BLAS calls. Because data transfer from the RAM to GPU suffers bandwidth limitations, we designed our implementation to transfer the A matrix and measurement vector at once.  The algorithm is similar
to the CPU algorithm except that we had to parallelize the code at an
additional level. The CPU portion of the code serves as the conductor, making
calls to GPU functions and directing the flow of the code while the GPU keeps all of the data and runs the computation.

% This section is Drew's
\section{Multiple Instance Implementations}

\subsection{Multiple Instance CPU Implementations}
%\subsection{Tuning the GEMV operation}
%When matrix-vector operations are called sequentially on the same data,
%the operation can be blocked so that each thread operates on the same block of data for each operation.
%For Single threaded MKL sgemv, Code took          0.0335823 msec per instance to run!
%For Automatic multi-threaded MKL sgemv, Code took 0.008179 msec per instance to run!
%For Manually multi-threaded MKL sgemv, Code took  0.0061276 msec per instance to run!

%\subsection{Tuning the Soft-Thresholding Operator}
%...Make sure there are no branches
%...Make sure there are no function calls
%...Make sure the loop gets automatically vectorized
%...Block the computation into work for different cpu cores
\subsection{Multiple Instance GPU Implementations}

% Describingn the sequence of GPU implementations Drew has tried...

\section{Benchmarks}
In this section, we present two sets of experiment to benchmark the performance
of the PALM algorithm on CPU and GPU platforms. 
The first set of experiments will compare implementations that solve a single
large instance of an $ell_1$ minimization problem on synthetic data to find at which
point the GPU implementation runs faster than CPU implementation.  
The second set of experiments will compare implementations that must solve
many instances of $\ell_1$ minimization.

To generate a meaningful comparison between different hardware architectures,
in all cases we compare the performance of hardware architectures on a
per-board basis:  For CPU implementations, the benchmark makes use of all of
the cores in as many cpu's are present.  For GPU implementations, the benchmark
makes the best use of the entire GPU chip (most GPU boards have a single GPU
chip).  Unless otherwise specified, all implementations utilize single
precision floating point datatypes.

\subsection{Single Large $\ell_1$ Problem, Synthetic Data}
%\subsection{PALM $\ell_1$ minimization benchmark on Synthetic Data}

@@

\subsection{Multiple Small $\ell_1$ Problems, Synthetic Data}

\subsection{Alignment Stage, Real Face Data}

>>>>

\section{Augmented Lagrangian Multiplier Algorithm}

@@

%In order to solve the alignment stage before recognition, we consider the $l_1$ problem: 
%\begin{equation} 
%\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
%\end{equation}
%which is solved in a similar fashion.

% A note on alignment l1...
% With a reshuffling of the order of operations and a change of the termination
% condition, the inner loop takes the form of two matrix-vector multiplications
% and a sequence of vector-vector operations.

\subsection{Alignment Stage L1 Computational Complexity} Any discussion of the speed of an
algorithm should discuss the comutational complexity of the algorithm.  For the
ALM algorithms considered in this paper, this is rather simple.  The
Matrix-vector computions are $O(mn)$, and the vector-vector operations are
$O(m)$, in terms of both computation cost and memory-bandwidth.  Since A is
very tall, the computation of $A^\dagger$, the pseudoinverse of $A$, is best
computed via $A^\dagger = (A^TA)^{-1} A^T$.  The complexity of computing $G =
A^T A$ is $O(nnm)$.  $G$ is sufficiently small that it can be inverted via
Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation
of $A$ finishes with a matrix multiplication, with cost $O(n^2 m)$.
Fortunately, $A^\dagger$ only needs to be computed once per instance of the
$\ell_1$ minimization problem.

@@

\section{Hardware Parallelism} 
In this section we give a brief overview of the hardware architectures used
in this paper.

We feel it is most meaningful to compare hardware architectures on a per-board basis,
and in a typical hardware configuration.  GPU's are most commonly shipped with a single
GPU chip per PCI-express card.  
%For multicore CPU architectures, a typical workstation
%will consist of

For CPU implementations, the benchmarks makes use of all of
the cores in as many cpu's are present.  For GPU implementations, the benchmark
makes the best use of the entire GPU chip (most GPU boards have a single GPU
chip).  Unless otherwise specified, all implementations utilize single
precision floating point datatypes.

@@

% This section is Drew's
\section{Iterative Alignment Stage Implementation}

\subsection{Iterative Alignment on the CPU}
%\subsection{Tuning the GEMV operation}
%When matrix-vector operations are called sequentially on the same data,
%the operation can be blocked so that each thread operates on the same block of data for each operation.
%For Single threaded MKL sgemv, Code took          0.0335823 msec per instance to run!
%For Automatic multi-threaded MKL sgemv, Code took 0.008179 msec per instance to run!
%For Manually multi-threaded MKL sgemv, Code took  0.0061276 msec per instance to run!

%\subsection{Tuning the Soft-Thresholding Operator}
%...Make sure there are no branches
%...Make sure there are no function calls
%...Make sure the loop gets automatically vectorized
%...Block the computation into work for different cpu cores
\subsection{Iterative Alignment on the GPU}
% Describingn the sequence of GPU implementations Drew has tried...

\subsection{Benchmark on Real Face Data}

\section{Recognition Stage Implementation}
% This section is Victor's

@@

\subsection{ GPU Implementation}
The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
CUBLAS are the BLAS libraries consisting of all single precision BLAS calls
which are provided by Nvidia and are optimized for GPU performance. In
addition, we coded kernel (GPU) methods for vector-vector operations to replace
operations that would have taken multiple BLAS calls. Because data transfer from the RAM to GPU suffers bandwidth limitations, we designed our implementation to transfer the A matrix and measurement vector at once.  The algorithm is similar
to the CPU algorithm except that we had to parallelize the code at an
additional level. The CPU portion of the code serves as the conductor, making
calls to GPU functions and directing the flow of the code while the GPU keeps all of the data and runs the computation.

\subsection{Benchmark on Synthetic Data}

@@

\section{Full Pipeline Benchmarks}
